rapr: Representative and Adequate Prioritisations in R

### 2016-06-24 code/rmarkdown/article.Rmd
+ assorted improvments to article + added in table s1 + added code to allow analysis and ms compilation on different machines

```{r, include=FALSE}
# load .rda file
try(session::restore.session('data/final/results.rda'))

# set options
options(error=NULL)
```

<!--  title page -->
Short running title: Representative and adequate prioritisations \newline
Word count: XXXXX \newline
Number of references: XXXX \newline
Number of figures: XXXX \newline
Number of tables: XXXX \newline
Number of textboxes: XXXX \newline
Number of equations: XXXX \newline
\newpage

# Introduction
<!-- conservation and reserve networks -->
The overarching aim of conservation is to maximise the long-term persistence of biodiversity [@r423; @r255]. To achieve this, conservation actions must preserve biodiversity patterns (eg. species, populations) and the processes that sustain them. One of the major tangible achievements of modern conservation has been the act of setting aside areas for preservation [@r440]. Reserve networks buffer species from threatening processes [eg. urbanisation; @r255] and set the stage for direct management interventions [eg. captive breeding and reintroduction programs; @r290]. However, the resources available for conservation action are limited, and so reserve networks must be sited in places that satisfy conservation objectives for minimum cost [@r255]. To achieve this, reserve selection is often formulated as an optimisation problem and then solved to identify cost-effective candidate reserve systems [prioritisations; @r255].

@@

\begin{align*}
T_s &= \text{amount target for feature $f$} \tag{1a}\\
%
\tau_{sa} &= \text{representation target for feature $f$ in attribute space $a$} \tag{1b}\\
%
B &= \text{boundary length modifier (BLM): penalise fragmented solutions} \tag{1c}\\
\end{align*}

@@

\begin{align*}
X_j
	&= \begin{cases}
		1, & \parbox{25em}{if planning unit $j$ is selected for conservation action} \tag{2a} \\
		0, & \parbox{25em}{otherwise} \\
	\end{cases} \\
%
Y_{fsij} &= \begin{cases}
		1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$. } \tag{2b} \\
		0, & \parbox{25em}{otherwise} \\
	\end{cases} \\
\end{align*}

@@

\begin{align*}
& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag{3a} \\
%
& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag{3b}\\
%
& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \left( \lambda_{fsi} d_{fsij} Y_{fsij} \right)^2}{\sum_{i=0}^{I-1} \left( \lambda_{fsi} \delta_{fsi} \right)^2 } \geq \tau_{fs}  & \forall & 0 \leq f \leq F-1, \tag{3c}\\
& & & & & 0 \leq s \leq S-1\\
%
& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag{3d}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1\\
%
& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag{3e}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J-1\\
%
& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag{3f}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1\\
%
\end{align*}

@@

## SIMULATION STUDY
### Methods
We simulated a hypothetical study area with planning units arranged in $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid (Fig. 2). We then simulated three species across this study area. The first species was simulated to represent a hyper-generalist--occurring in all planning units with equal probability (Fig. 2a). This species' distribution was based on a uniform distribution (eqn 8a). The second species was simulated to represent a species with an idealised distribution (Fig. 2b). It has a core range area, and is less likely to be found in areas that are distant from the core area. This species' distribution was simulated using the probability density function of a single multivariate normal distribution (eqn 8b). The third species represents a species with two distinct populations (Fig. 2c). This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 8c). The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 8a--8c. We used a geographic attribute space to showcase the behavior of the problem. For each species, demand points were generated by calculating the centroids of each planning units and their weights were set as the probability that the planning units were occupied.

\begin{align*}
& P (\text{uniform species} | x, y) &= 0.1 & \tag{8a}\\
& P (\text{normal species} | x, y) &= XXXX & \tag{8b}\\
& P (\text{uniform species} | x, y) &= XXX & \tag{8c}\\
\end{align*}

@@

## CASE STUDY 1
### Methods
Here we investigated how space-based targets can be used in a multi-species planning context to generate a prioritisation that sufficiently preserves the realised niche for several species. We selected Queensland, Australia as the study area. This region is ideal for exploring the potential of niche-based targets because it contains a broad range of different habitats. We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subject them to a principal components analysis (using ArcMap 10.3.1). We used the first `r as.character(as.english(cs1.params.LST[[MODE]]$n.pc.layers))` principle components (cumulatively explained `r round(cs1.pca.DF[[4]][cs1.params.LST[[MODE]]$n.pc.layers],1)` % of the total variation) to characterise the environmental variation across the study area (Fig. 4).

@@

### Results
Generally, the solution generated using amount targets preserved a representative sample of each the `r length(cs1.params.LST[[MODE]]$species.names)` bird species' niches--with exception to `r ifelse(length(amount.not.represented.species.names)>1, as.character(as.english(length(amount.represented.species.names))), 'a single')` bird species. Specifically, this solution sampled over `r cs1.params.LST[[MODE]]$space.target*100` % of the realised niche of `r parsed.representative.space.held.names`. Yet it failed to achieve this for `r parsed.not.representative.space.held.names`.

@@

>>>>

```{r, include=FALSE}
## initialization
# load .rda file
try(session::restore.session('data/final/results.rda'))

# set options
options(error=NULL)

## preliminary processing
for (i in seq_len(nlayers(cs1.bioclim.RAST)))
	cs1.bioclim.RAST@layers[[i]]@file@name  <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'rapr-manuscript')[[1]][[2]])

```

<!--  title page -->
Short running title: Representative and adequate prioritisations \newline
Word count: 322 / 350 \newline
Number of references: 34 \newline
Number of figures: 8 \newline
Number of tables: 0 \newline
Number of textboxes: 0 \newline
Number of equations: 8 \newline
\newpage

# Introduction
<!-- conservation and reserve networks -->
The overarching aim of conservation is to maximise the long-term persistence of biodiversity [@r423; r255]. To achieve this, conservation actions must preserve biodiversity patterns (eg. species, populations) and the processes that sustain them. One of the major tangible achievements of modern conservation has been the act of setting aside areas for preservation [@r440]. Reserve networks buffer species from threatening processes [eg. urbanisation; @r255] and set the stage for direct management interventions [eg. captive breeding and reintroduction programs; @r290]. However, the resources available for conservation action are limited, and so reserve networks must be sited in places that satisfy conservation objectives for minimum cost [@r255]. To achieve this, reserve selection is often formulated as an optimisation problem and then solved to identify cost-effective candidate reserve systems [prioritisations; @r255].

@@

\begin{align*}
T_s &= \text{amount target for feature $f$} \tag*{eqn 1a}\\
%
\tau_{sa} &= \text{representation target for feature $f$ in attribute space $a$} \tag*{eqn 1b}\\
%
B &= \text{boundary length modifier (BLM): penalise fragmented solutions} \tag*{eqn 1c}\\
\end{align*}

@@

\begin{align*}
X_j
	&= \begin{cases}
		1, & \parbox{25em}{if planning unit $j$ is selected for conservation action} \tag*{eqn 2a} \\
		0, & \parbox{25em}{otherwise} \\
	\end{cases} \\
%
Y_{fsij} &= \begin{cases}
		1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$. } \tag*{eqn 2b} \\
		0, & \parbox{25em}{otherwise} \\
	\end{cases} \\
\end{align*}

@@

\begin{align*}
& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag*{eqn 3a} \\
%
& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag*{eqn 3b}\\
%
& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \left( \lambda_{fsi} d_{fsij} Y_{fsij} \right)^2}{\sum_{i=0}^{I-1} \left( \lambda_{fsi} \delta_{fsi} \right)^2 } \geq \tau_{fs}  & \forall & 0 \leq f \leq F-1, \tag*{eqn 3c}\\
& & & & & 0 \leq s \leq S-1\\
%
& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 3d}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1\\
%
& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 3e}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J-1\\
%
& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3f}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1\\
%
\end{align*}

@@

## SIMULATION STUDY
### Methods
We simulated a hypothetical study area with planning units arranged in $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid (Fig. 2). We then simulated three species across this study area. The first species was simulated to represent a hyper-generalist--occurring in all planning units with equal probability (Fig. 2a). This species' distribution was based on a uniform distribution (eqn 8a). The second species was simulated to represent a species with an idealised distribution (Fig. 2b). It has a core range area, and is less likely to be found in areas that are distant from the core area. This species' distribution was simulated using the probability density function of a single multivariate normal distribution (eqn 8c). The third species represents a species with two distinct populations (Fig. 2c). This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 8d). The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 8a--8d. We used a geographic attribute space to showcase the behavior of the problem. For each species, demand points were generated by calculating the centroids of each planning units and their weights were set as the probability that the planning units were occupied.

\begin{align*}
\\
P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.1 \tag*{eqn 8a} \\
%
f \left(z, \mu, \Sigma \right) &= \left(2 \pi \right) ^{-1} | \Sigma | e ^{- \frac{1}{2} \left(z - \mu \right)' \Sigma ^{-1} \left(z - \mu \right)} \tag*{eqn 8b} \\
%
P \left( \text{normally distributed species} | \left( x , y \right)  \right) &= \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 8c} \\
%
P \left( \text{bimodally distributed species} | \left(  x,y \right) \right) &= \text{Max} \begin{cases}
f \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
\end{cases} \tag*{eqn 8d} \\
%
\end{align*}

@@

## CASE STUDY 1
### Methods
Here we investigated how space-based targets can be used in a multi-species planning context to generate a prioritisation that sufficiently preserves the realised niche for several species. By preserving the populations in suitable habitats with different with environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience against environmental change [@r63]. We selected Queensland, Australia as the study area. This region is ideal for exploring the potential of niche-based targets because it contains a broad range of different habitats. We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subject them to a principal components analysis (using ArcMap 10.3.1). We used the first `r as.character(as.english(cs1.params.LST[[MODE]]$n.pc.layers))` principle components (cumulatively explained `r round(cs1.pca.DF[[4]][cs1.params.LST[[MODE]]$n.pc.layers],1)` % of the total variation) to characterise the environmental variation across the study area (Fig. 4).

@@

### Results
Generally, the solution generated using amount targets preserved a representative sample of each the `r length(cs1.params.LST[[MODE]]$species.names)` bird species' niches--with exception to `r ifelse(length(amount.not.represented.species.names)>1, as.character(as.english(length(amount.represented.species.names))), 'a single')` bird species. Specifically, this solution sampled over `r cs1.params.LST[[MODE]]$space.target*100` % of the realised niche of `r parsed.representative.space.held.names`. Yet it failed to achieve this for `r parsed.not.representative.space.held.names`. This result suggests that prioritisations generated using conventional methods may preserve a representative sample of most species realised niches (Figure S1). However, despite this, there may yet be species for which only a small fraction of their realised niche is preserved. By explictly using space targets, conservation planners can generate prioritisations that guaranteeably preserve a representative sample of species' niches.

@@

\bibliography{references}

### 2016-06-24 code/rmarkdown/figures.Rmd
+ assorted improvments to article + added in table s1 + added code to allow analysis and ms compilation on different machines

```{r, include=FALSE}
# load .rda file
try(session::restore.session('data/final/results.rda'))

@@

# prelimary processing
sim.pus.FPLY <- fortify(sim.pus)
aus.FPLY <- fortify(aus.PLY)
cs1.pus.FPLY <- fortify(cs1.pus)
cs2.grid.FPLY <- fortify(cs2.grid.PLY)
cs2.grid.sub.FPLY <- fortify(cs2.grid.sub.PLY)

>>>>

```{r, include=FALSE}
## initialization
# load .rda file
try(session::restore.session('data/final/results.rda'))

@@

## prelimary processing
# redefine file paths for rasters
for (i in seq_len(nlayers(cs1.bioclim.RAST)))
	cs1.bioclim.RAST@layers[[i]]@file@name  <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'rapr-manuscript')[[1]][[2]])

# fortify polygons
sim.pus.FPLY <- fortify(sim.pus)
aus.FPLY <- fortify(aus.PLY)
cs1.pus.FPLY <- fortify(cs1.pus)
cs2.grid.FPLY <- fortify(cs2.grid.PLY)
cs2.grid.sub.FPLY <- fortify(cs2.grid.sub.PLY)

### 2016-06-24 code/rmarkdown/supporting-information.Rmd
+ assorted improvments to article + added in table s1 + added code to allow analysis and ms compilation on different machines

```{r, include=FALSE}
# load .rda file
try(session::restore.session('data/final/results.rda'))

# set options
options(error=NULL)
```

@@

## References

>>>>

```{r, include=FALSE}
## initialization
# load .rda file
try(session::restore.session('data/final/results.rda'))

# set options
options(error=NULL)

## prelimary processing
# redefine file paths for rasters
for (i in seq_len(nlayers(cs1.bioclim.RAST)))
	cs1.bioclim.RAST@layers[[i]]@file@name  <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'rapr-manuscript')[[1]][[2]])

```

@@

```{r, echo=FALSE, results='asis'}
# make table
term.CHR <- c(
	'$F$', 'set of biodiversity features (indexed by $f$)',
	'$J$', 'set of planning units (indexed by $j$)',
	'$q_{fj}$', 'probability of feature $f$ occupying planning unit $j$',
	'$e_{jk}$', 'shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrise exposed edges with no neighbours.',
	'$S$', 'set of attribute spaces (indexed by $S$)',
	'$I_{fsi}$', 'set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$',
	'$\\lambda_{fsi}$', 'set of weights for demand point $i$ for feature $f$ in attribute space $s$.',
	'$d_{fsij}$', 'distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$.',
	'$\\delta_{fsi}$', 'the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$.',
	'$T_f$', 'amount target for feature $f$.',
	'$\\tau_{fs}$', 'space-based target for feature $f$ in attribute space $s$.',
	'$X_j$', 'binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0).',
	'$Y_{fsij}$', 'binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determing the amount of the attribute space sampled by the selected planning units.'
)
term.DF <- data.frame(
	Symbol=term.CHR[seq(1, length(term.CHR), 2)],
	Description=term.CHR[seq(2, length(term.CHR), 2)]
)
term.DF$Description <- paste0('\\parbox[t][][t]{15cm}{', term.DF$Description, '}')

# render table
curr.file <- tempfile(fileext='.tex')
tmp <- latex(
		term.DF,
		file=curr.file, 
		digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
		colheads=c('Symbol', 'Description'), col.just=c('c', 'l'),
		caption='Symbols and descriptions of terms used in the formulation of the unreliable representative and adequate prioritisation (URAP) problem.'
	)
readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
	cat()
```

# References
